\begin{abstract}
\noindent
XLM-R is a strong multilingual masked language model covering 100 languages, but it
over-fragments tokens of unseen tail languages (tokens/word = 2.204) and fails to align
their lexical and morphological features in the representation space.
Glot500 addresses this by extending XLM-R's vocabulary with a mixed-corpus SentencePiece
tokenizer and continuing pretraining via masked language modeling (MLM), and reports
improved tail-language representations through related-language help.
A key step Glot500 leaves open is how to initialize the new token embedding rows.

We investigate this choice with a controlled 5-way ablation on \texttt{xlm-roberta-base}.
The tokenizer is fixed to Glot500-style SentencePiece append, adding 116,664 new pieces,
and only the initialization of those new rows varies across five methods:
\rand, \meanm, \fvt\ (Fast Vocabulary Transfer),
\wfvt\ (length-weighted FVT), and \famm\ (family-aware mean).
We train on 92 XLM-R-seen replay languages and 7 XLM-R-unseen target languages (Target7)
for 50K steps under identical corpus, objective, and evaluation protocol.
Evaluation follows the Glot500 downstream suite:
pseudoperplexity (PPPL), Tatoeba and Bible sentence retrieval, NER, roundtrip alignment,
and text classification.

We observe three findings.
(1) The extended tokenizer reduces Target7 tokens/word from 2.204 to 1.592 (27.75\% reduction),
confirming that tail-language fragmentation is alleviated.
(2) Initialization alone changes 50K convergence loss and downstream scores:
\famm\ achieves the lowest target PPPL (12.3) and highest Tatoeba tail accuracy (36.1),
while \wfvt\ is strongest on training loss and roundtrip alignment.
FVT-family methods lead at 10K steps but are overtaken by \wfvt\ and \famm\ by 30--50K,
showing that the benefit of structured initialization emerges at convergence, not at early checkpoints.
(3) The sentence representation space exhibits language-identity and family/typology clustering,
and languages with strong family coverage in the head set benefit most from \famm.
In short, embedding initialization is a consequential design choice for
low-resource vocabulary adaptation, and corpus-provenance or family-prior initialization
outperforms random initialization in the target regime.
\end{abstract}
