\section{Related Work}

\subsection{Glot500 and Horizontal Language Scaling}

Glot500 \citep{imani2023glot500} is the horizontal-scaling approach most directly related to
this work.
It treats languages seen during XLM-R training as \emph{head languages} and all others as
\emph{tail languages}, and scales the model to cover 511 language-scripts.

\paragraph{Corpus collection.}
Glot500-c aggregates text from religious texts (Bible translations), news articles, scientific
papers, and web crawls.
Data are organized by language-script unit and filtered with chunk-level filters
(character repetition, word repetition, special characters, short texts, duplicates)
and corpus-level filters (language--script mismatch, perplexity divergence against a 3-gram
character LM).
Only language-scripts with at least 30,000 training sentences are retained.

\paragraph{Tokenization.}
Glot500 trains a SentencePiece unigram tokenizer \citep{kudo2018unigram,kudo2018sentencepiece}
on Glot500-c with temperature sampling ($\alpha=0.3$) and merges the resulting vocabulary
into XLM-R's existing SentencePiece model by appending pieces not already present.
Byte fallback prevents unknown tokens.

\paragraph{Continued pretraining.}
Glot500-m is produced by continued MLM pretraining of XLM-R-base on Glot500-c.
Language sampling uses $\alpha=0.3$ to prevent high-resource head languages from dominating.

\paragraph{Key findings.}
Glot500 reports that performance depends jointly on corpus size, script coverage, related-language
support, and model capacity.
Notably, having related languages in the training set improves tail-language representations
(\emph{related-language help}).
Our \famm\ initialization directly encodes this prior into the new embedding rows.

\subsection{Two Vocabulary Extension Approaches}

Two vocabulary extension designs appear in the literature.
The \emph{mixed-corpus} approach (Glot500-style) trains a new tokenizer on a corpus of
both seen and unseen languages and appends all new pieces to the source vocabulary,
resulting in a large number of new tokens (116,664 in our experiment).
The \emph{target-only} approach \citep{yamaguchi2025vocab} trains an auxiliary tokenizer
on a small amount of target-language text ($\approx$0.01~GB) and appends only the target-specific
pieces, keeping the vocabulary increment small.
This report fixes the main axis to Glot500-style mixed-corpus injection and compares only
initialization; the Yamaguchi-style approach is a separate axis not varied here.

\subsection{Embedding Initialization for New Tokens}

Placing new vocabulary rows in the correct region of an existing embedding space has been
studied under the name \emph{vocabulary transfer}.

\textbf{FVT} \citep{gee2022fvt} decomposes each new token's surface form with the source
tokenizer and initializes the row as the mean of the constituent source embeddings.
Our \fvt\ implements this directly.

\textbf{WECHSEL} \citep{minixhofer2022wechsel} computes similarity between new tokens and
source tokens using static multilingual word embeddings and initializes new rows as
a weighted combination of similar source rows.

\textbf{FOCUS} \citep{dobler2023focus} uses anchor tokens that overlap between source and
target vocabularies to express new tokens as combinations of anchors.

Both WECHSEL and FOCUS demonstrate that reusing semantic or surface similarity beats
random initialization---the same motivation behind \famm, which reuses typological similarity.
We do not implement WECHSEL or FOCUS because they require external static embeddings or
separate anchor computation; we cite them as background.

\textbf{Mean initialization} \citep{hewitt2021newembeddings} places new rows near the
centroid of existing embeddings to reduce early instability, which is our \meanm\ baseline.

\subsection{Evaluation Protocol}

We follow the Glot500 downstream suite: PPPL \citep{salazar2020pppl}, Tatoeba and Bible
sentence retrieval \citep{artetxe2019tatoeba}, NER on WikiANN \citep{pan2017wikiann},
roundtrip word alignment \citep{dufter2018roundtrip}, and text classification on
Taxi1500 \citep{ma2023taxi1500}.
Since labeled training data does not exist for tail languages, all supervised tasks
fine-tune on English and evaluate zero-shot on target languages, following the
XTREME protocol \citep{hu2020xtreme}.
This setup ensures that cross-lingual alignment---the quantity that initialization
directly affects---is the only pathway to tail-language performance.
