\section{Data}

\subsection{Language Scope and Data Sources}

Our corpus universe consists of 99 language-scripts drawn from Glot500-c:
92 \emph{head} language-scripts that XLM-R was trained on, and 7 \emph{tail}
language-scripts that XLM-R did not see (Target7).
Both groups use Glot500-c as their text source, accessed as HuggingFace Arrow datasets
(\texttt{\{lang\}\_\{script\}}, \texttt{train} split, \texttt{text} column).

Head languages serve as replay data: including them in continued pretraining prevents
catastrophic forgetting of XLM-R's existing representations.
Target7 languages are the evaluation targets.
They satisfy four selection criteria:
(i) absent from XLM-R training,
(ii) present in Glot500-c,
(iii) at least 30,000 training sentences (\texttt{new\_length} $\geq$ 30,000),
and (iv) at least one downstream evaluation dataset available.

\begin{table}[h]\centering\small
\caption{Target7 languages with family, region, corpus size, and covered downstream tasks.}
\label{tab:target7}
\begin{tabular}{@{}llllrl@{}}
\toprule
language\_script & Name & Family & Region & new\_length & Tasks \\
\midrule
\texttt{dtp\_Latn} & Kadazan Dusun & Austronesian & SE Asia & 48,468 & Tatoeba, Bible, Roundtrip, Taxi, Embed. \\
\texttt{xav\_Latn} & Xavante & Macro-Je & S America & 31,765 & Bible, Roundtrip, Taxi \\
\texttt{bam\_Latn} & Bambara & Mande & W Africa & 32,150 & Bible, Roundtrip, Taxi \\
\texttt{csb\_Latn} & Kashubian & Slavic & Europe & 33,743 & Tatoeba, NER, Embed. \\
\texttt{ile\_Latn} & Interlingue & Constructed & Europe & 40,984 & Tatoeba, Embed. \\
\texttt{lij\_Latn} & Ligurian & Romance & Europe & 42,447 & NER \\
\texttt{fur\_Latn} & Friulian & Romance & Europe & 30,052 & NER \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Task-Centered Language Selection}

The seven languages were selected in reverse from the evaluation tasks, not arbitrarily.
Most genuinely low-resource languages lack labeled downstream data, making it impossible to
complete the tokenizer $\to$ training $\to$ evaluation pipeline end to end.
We therefore required at least three evaluable target languages per downstream task,
as shown in Table~\ref{tab:taskcov}.
Geographic and typological spread (SE Asia, South America, West Africa, Europe) was
also considered, but traceability---being able to connect a specific score to a specific
language---was the primary criterion.

\begin{table}[h]\centering\small
\caption{Task coverage: three Target7 languages per downstream task.}
\label{tab:taskcov}
\begin{tabular}{@{}lll@{}}
\toprule
Task & Target languages & Scale \\
\midrule
Tatoeba retrieval & \texttt{dtp}, \texttt{ile}, \texttt{csb} & 2,253 sentence pairs \\
Bible retrieval & \texttt{dtp}, \texttt{xav}, \texttt{bam} & 23,238 sentence pairs \\
Roundtrip alignment & \texttt{dtp}, \texttt{xav}, \texttt{bam} & 22,669 sentence groups \\
NER & \texttt{csb}, \texttt{lij}, \texttt{fur} & 300 sentences (100/lang.) \\
Embedding similarity & \texttt{csb}, \texttt{dtp}, \texttt{ile} & 2,100 sentence pairs \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Training Corpus and Temperature Sampling}

The MLM training corpus mixes head replay and Target7 text.
To prevent high-resource head languages from dominating the vocabulary and gradient signal,
we apply temperature sampling with $\alpha=0.3$:
\[
P_i \propto n_i^{\alpha}, \quad \alpha = 0.3,
\]
where $n_i$ is the corpus size of language-script $i$.
At $\alpha=1$, a head language with one million sentences would be sampled 1,000$\times$
more often than a tail language with 1,000 sentences;
at $\alpha=0.3$ the ratio shrinks to approximately 8:1.
This downweighting applies both to tokenizer training and to the actual MLM corpus.

\begin{table}[h]\centering\small
\caption{Training and evaluation corpus sizes.}
\label{tab:corpus}
\begin{tabular}{@{}lr@{}}
\toprule
Item & Count \\
\midrule
Source raw sentences & 1,025,895,043 \\
MLM train --- seen/head & 4,147,176 \\
MLM train --- Target7 & 335,083 \\
MLM train --- total & 4,482,259 \\
\midrule
Eval --- PPPL & 700 \\
Eval --- Tatoeba & 2,253 \\
Eval --- Bible & 23,238 \\
Eval --- Roundtrip & 22,669 \\
Eval --- NER & 300 \\
\bottomrule
\end{tabular}
\end{table}
