\section{Experiments}

All five initialization methods share the same tokenizer, corpus, MLM objective, training budget,
and evaluation runner.
Score differences therefore reflect only the interaction between initialization and
training dynamics.

\paragraph{Common sentence embedding.}
Retrieval and similarity tasks use the Glot500 retrieval-style encoder:
hidden states from layer 7 (0-indexed) are mean-pooled over non-special tokens,
then L2-normalized.
For centered cosine analysis, the global sentence-vector mean is subtracted before normalization.

\subsection{Pseudoperplexity (PPPL \dn)}

Following \citet{salazar2020pppl}, we compute pseudoperplexity over the held-out target
sentence pool.
PPPL masks each token position in turn and accumulates the negative log-likelihood of
the true token:
\[
\mathrm{PPPL}(x) = \exp\!\left(\frac{1}{n}\sum_{t=1}^{n} -\log p_\theta(x_t \mid x_{\setminus t})\right).
\]
Lower is better.
We use 100 sentences per Target7 language (700 total).
Because all five methods use the same tokenizer, PPPL differences across methods
cannot arise from tokenizer fertility; they reflect model quality alone.

\subsection{Sentence Retrieval --- Tatoeba and Bible (\up)}

Following \citet{hu2020xtreme} and \citet{artetxe2019tatoeba}, we embed English and target
sentences and retrieve the top-10 nearest neighbors by cosine similarity.
A query is a hit if the correct translation appears in the top 10 (Acc@10).
Tatoeba provides up to 1,000 English-aligned sentence pairs per language;
Bible provides $\sim$7,000--8,000 verse-aligned pairs per language, making it
substantially harder (see Appendix~A for analysis of the Bible floor).

\subsection{NER (\up)}

We follow the XTREME NER protocol \citep{hu2020xtreme} using WikiANN \citep{pan2017wikiann}.
A token-classification head is attached to the encoder and fine-tuned on English WikiANN
for 10 epochs (LR $2\times10^{-5}$, batch $8\times4$, max length 256).
Zero-shot transfer to the three NER-evaluable Target7 languages (\texttt{csb, lij, fur})
is evaluated with seqeval entity-level F1 on BIO-tagged sequences
(\texttt{O, B-PER, I-PER, B-ORG, I-ORG, B-LOC, I-LOC}).

\subsection{Roundtrip Alignment (\up)}

Following \citet{dufter2018roundtrip}, we assess representation quality without gold labels
by computing word alignment through a language cycle and measuring the fraction of source words
that return to themselves.
We use SimAlign \citep{jalili2020simalign} with layer-7 embeddings on the Bible corpus.
Each experiment uses five random sets of three intermediate language-scripts and reports the mean.

\subsection{Text Classification --- Taxi1500 (\up)}

We use Taxi1500 \citep{ma2023taxi1500} for 6-class text classification.
A sequence-classification head is fine-tuned on English for 30 epochs
(LR $2\times10^{-5}$, effective batch 16).
Because the local Taxi1500 materialization covers only English, we report only the
English head score; target-language transfer is marked \texttt{not\_applicable}.

\paragraph{Alignment with Glot500 protocol.}
NER learning rate 2e-5 with Adam and English-only fine-tuning matches the Glot500 paper.
Text classification effective batch 16 matches Glot500.
Retrieval, roundtrip, and PPPL involve no fine-tuning, so initialization and MLM training
are the sole sources of score differences.
