\section{Limitations and Conclusion}

\subsection{Limitations}

This study is a controlled budget ablation; its conclusions are bounded by the following scope.

\begin{itemize}
  \item \textbf{Scale.} The experiment uses 92 head + 7 tail language-scripts, not the 511
        language-scripts of full Glot500.
        Findings may not generalize to a broader language set.

  \item \textbf{Script diversity.} All Target7 languages use Latin script.
        We make no claim about script diversity or about languages with non-Latin scripts.

  \item \textbf{Family coverage asymmetry.} \famm's advantage is concentrated in
        \texttt{dtp} (Austronesian), which has five head-language family members.
        Languages with no usable family prior (\texttt{xav}) show no advantage from \famm.
        Whether family-aware initialization generalizes beyond well-covered families
        requires a broader head set.

  \item \textbf{Single run.} \wfvt\ and \famm\ are each a single run.
        Multi-seed replication is needed to confirm that their advantages are stable.

  \item \textbf{Training budget.} The 50K-step window is sufficient to observe crossover
        and convergence, but longer training may change rankings.

  \item \textbf{Task coverage asymmetry.} NER covers three tail languages; Bible retrieval
        and roundtrip cover a different three.
        The aggregate tail scores therefore reflect different language subsets per metric.

  \item \textbf{Text classification.} The local Taxi1500 materialization covers only English,
        so Text(EN) does not measure target-language transfer.
\end{itemize}

\subsection{Conclusion}

We applied Glot500-style vocabulary extension to \texttt{xlm-roberta-base} and compared
five new-token embedding initialization methods in a controlled 50K-step continued MLM
pretraining experiment.
Our findings are:

\begin{enumerate}
  \item \textbf{Tokenizer fertility.} The extended tokenizer reduces Target7 tokens/word
        by 27.75\% (2.204 $\to$ 1.592), confirming that tail-language fragmentation is
        substantially reduced.

  \item \textbf{Initialization affects convergence loss.}
        Under identical tokenizer, corpus, and objective, the five methods converge to
        distinct MLM loss values: \wfvt\ 2.73 $<$ \fvt\ 2.76 $<$ \famm\ 2.91
        $<$ \rand\ 3.12 $<$ \meanm\ 3.27.
        \meanm\ is worse than \rand, showing that collapsing all new rows to a single
        centroid is harmful.

  \item \textbf{Family-aware initialization leads at convergence.}
        \famm\ achieves the lowest target PPPL (12.3) and highest Tatoeba tail accuracy (36.1)
        at 50K, while \wfvt\ is the most balanced method across all metrics.
        The benefit is not visible at 10K---\fvt\ leads early---but \wfvt\ and \famm\
        overtake \fvt\ by 30K and maintain the lead through 50K.
        Evaluating initialization from early diagnostics alone would yield the wrong ranking.

  \item \textbf{Family structure in the representation space explains the advantage.}
        Centered cosine analysis shows that the encoder groups languages by typological family,
        not merely by script, and that this structure forms by 10K steps and is stable
        thereafter.
        \famm\ places new token rows in the correct family region of the embedding space
        from the start, shortcutting the alignment process that \rand\ and \meanm\
        must complete from scratch during training.
\end{enumerate}

In low-resource vocabulary adaptation, embedding initialization is a consequential design
choice.
Reusing corpus-provenance information (\fvt, \wfvt) outperforms random initialization,
and explicitly encoding language-family structure (\famm) provides additional benefit
for languages with adequate family coverage in the head set.
Future work should validate these findings on non-Latin scripts, a wider family distribution,
and longer training budgets, and should explore combining surface-form and family priors
based on provenance confidence.
